---
title: "Analysis of the Stock market behaviour on COVID-19 Pandemic"
author: "cutiepootie"
date: "December, 2020"
output:
  pdf_document: default
  html_document: default
---
*Code and data supporting this analysis is available at: https://github.com/------*
```{r setup, include=FALSE, python.reticulate = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#install.packages("quantmod")
#library("quantmod")
#install.packages("tidyquant")
library("tidyquant")
#why isnt it running
library(kableExtra)
library(knitr)
library(tidyverse)
```
# Abstract  

# Keywords  
Difference in differences, 
control group(stocks that do not get affected by covid-19), treatment group(stocks get affected by covid-19)

# Introduction
The world has been overturned by an unexpected virus COVID-19, and it seems that the stock market has been under the spotlight since the Covid-19 pandemic. The 2020 Stock market crash, was a major global stock market crash that began around March since the world has officially turned into a pandemic, and was in an emergency. The spread of the virus caused the Dow Jones to crash by almost 40%. Compared to other historically significant stock market crashes such as Black Monday of 1987(-31%), the 2000s Recession(-34%), and the Great Recession of 2008(-49%), 2020 Covid stock market crash shows a very significant economic recession. However, after a few months, the stock market recovered surprisingly fast, And depending on the industry of the securities, some stocks have been showing amazing growth, and some stocks are still in the process of recovery.  

Throughout the report, we are going to analyze some market index to determine which industries have shown the most correlation to the pandemic, and analyze the causal effect using the difference in differences between two stocks, one being assumed as a "treatment" group, where they get negatively or positively affected by the pandemic. The other stock would be in a "control" group, where the stock hasn't been affected by coronavirus. The difference is differences are measured by two different variables, control and treatment, and by first calculating the difference in first and second time periods, and then subtracting the average change in the control group from the average change in the treatment group(3).  

Three stock data will be used to investigate the effects of the COVID-19 pandemic on the stock market, stock a positive effect, negative effect, and no effect. In the Methodology section, I describe the data and the model that was used to see the time series trend of each stock's price. Results of the Difference in differences are provided in the Results section.  

# Back up options
I have some other options if this doesn't come out well. 

## Option A
Drug uses - maybe causual inference on drug uses/life satisfaction(i would have to simulate data for this) or study hours, or GPA
and regression analysis along with it.   

## Option B
Using GSS data - similar approach as PS2, but using completely different factors for regression analysis. 
I could probably also do a causal inference with the data.   

I'm really being indecisive here, so I would appreciate a lot(and I mean, A LOT) if you guys help me choosing one topic <3 Thank you!

# Methodology

## Data
Data API is from YahooFinance, and Major stocks that are going to be used for the causal inference are downloaded. Each data contains opening price, closing price, highest price, lowest price, and adjusted price for each date. 
```{r, echo = F, warning = F, include = F}
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)
# Downloading Apple price using quantmod

getSymbols("AAPL", from = '2020-01-01',
           to = "2020-12-01",warnings = FALSE,
           auto.assign = TRUE)

getSymbols("NDAQ", from = '2020-01-01',
           to = "2020-12-01",warnings = FALSE,
           auto.assign = TRUE)

getSymbols("MRNA", from = '2020-01-01',
           to = "2020-07-01",warnings = FALSE,
           auto.assign = TRUE)
Moderna.pre <- MRNA

getSymbols("AC.TO", from = '2020-01-01',
           to = "2020-07-01",warnings = FALSE,
           auto.assign = TRUE)
AirCanada.pre <- AC.TO

getSymbols("MRNA", from = '2020-09-01',
           to = "2020-12-06",warnings = FALSE,
           auto.assign = TRUE)

getSymbols("AC.TO", from = '2020-09-01',
           to = "2020-12-06",warnings = FALSE,
           auto.assign = TRUE)

Moderna.post <- MRNA
AirCanada.post <- AC.TO
```

```{r echo = F}

# Moderna pre & post
kable(head(Moderna.pre), caption = "Moderna pre Pandemic")%>% kable_styling(latex_options = "HOLD_position")
kable(head(Moderna.post), caption = "Treatment group (Moderna)")%>% kable_styling(latex_options = "HOLD_position")

#Air Canada pre & post 
kable(head(AirCanada.pre), caption = "Air Canada pre Pandemic")%>% kable_styling(latex_options = "HOLD_position")
kable(head(AirCanada.post), caption = "Treatment group (Air Canada)")%>% kable_styling(latex_options = "HOLD_position")
```




## Model
```{r}
plot(MRNA)
plot(AC.TO)
```

# Results 

# Discussion


# Reference
1. Stock market API: Coding Finance -. (2018, April 02). Retrieved December 09, 2020, from https://www.codingfinance.com/post/2018-03-27-download-price/

2. 2020 Stock Market Crash: Team, T. (2020, March 13). Market Crashes Compared: -28% Coronavirus Crash Vs. 4 Historic Market Crashes. Retrieved December 09, 2020, from https://www.forbes.com/sites/greatspeculations/2020/03/13/market-crashes-compared28-coronavirus-crash-vs-4-historic-market-crashes/?sh=55d01ac74ee8

3. DID: https://www.statisticshowto.com/difference-in-differences/

# Appendix

